### 2015-12-07 report1_high_level/manuscript.Rmd
A tight edit of the draft manuscript

---
title: 'Open Software - Restricted Data: A case study.'
author:  
- name: Ivan C. Hanigan
  affilnum: 1
  email: ivan.hanigan@anu.edu.au  
- name: David Fisher
  affilnum: 2
- name: Steven McEachern
  affilnum: 3
affiliation:
- affilnum: 1
  affil: National Centre for Epidemiology and Population Health (ANU) 
- affilnum: 2
  affil: Information Technology Services (ANU)
- affilnum: 3
  affil: Australian Data Archives (ANU)
header-includes:
  - \usepackage{graphicx}
  - \usepackage{url}   
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    template: components/manuscript.latex
  html_document: null
  word_document: null
fontsize: 11pt
capsize: normalsize
csl: meemodified.csl
documentclass: article
classoption: a4paper
spacing: singlespacing
linenumbers: no
bibliography: references.bib
abstract: no
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Example Manuscript}
-->
```{r, eval = F, echo = F}
setwd("~/projects/opensoftware-restricteddata.github.com/report1_high_level/")
library(rmarkdown)
library(knitr)
library(knitcitations)
library(bibtex)
cleanbib()
#rm("bib")
#options("cite_format"="pandoc")
cite_options(citation_format = "pandoc", check.entries=FALSE)

dir()
render("manuscript.Rmd", "pdf_document")
browseURL("manuscript.pdf")
```
```{r, echo = F, results = 'hide'}
# load
if(!exists("bib")){
bib <- read.bibtex("~/references/library.bib")

for(bibkey in c("SarathiBiswas2012",
  "Mcmichael2002a", "Gelman2013"
)){
bib[[bibkey]]$url <- gsub("\\{\\\\_\\}","_", bib[[bibkey]]$url)
bib[[bibkey]]$url <- gsub("\\{~\\}","~", bib[[bibkey]]$url)
}

}
```
```{r setup, include=FALSE, echo=FALSE}
#Put whatever you normally put in a setup chunk.
#I usually at least include:
#devtools::install_github("manuscriptPackage","jhollist")
#library("manuscriptPackage")
#Didn't do that here to expedite building of the example vignette
library("knitr")

opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=5)

# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
#Figure captions are handled by LaTeX

knit_hooks$set(tab.cap = function(before, options, envir) {
                  if(!before) { 
                    paste('\n\n:', options$tab.cap, sep='') 
                  }
                })
default_output_hook = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (is.null(options$tab.cap) == FALSE) {
    x
  } else
    default_output_hook(x,options)
})
```

```{r analysis , include=FALSE, echo=FALSE, cache=FALSE}
#All analysis in here, that way all bits of the paper have access to the final objects
#Place tables and figures and numerical results where they need to go.
```

<!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. -->

Draft: \today

\singlespace

\vspace{2mm}\hrule
\paragraph*{Background:} This essay was written to accompany the material presented as a
speedtalk and poster at the National Climate Change Adaptation
Research Facility Conference 'Climate Adaptation knowledge and
partnership', June 2013, Sydney.  The poster and slideshow are both available to download from this website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/)

\paragraph*{Methods:} The paper reports on a project to build tools and procedures for enhancing open and transparent analysis of restricted datasets.
Some datasets such as suicide or climate
change scenarios need to be accessed in a restricted way.  On the
other hand scientists need to make their methods, models and
assumptions transparent and available for scientific debate even
though the datasets may require authorisation to access.

\paragraph*{Results:} We built a secure server/client computational
environment for using open software with restricted data.  We
demonstrate the use of this system using drought and suicide as a case study.  
We describe the potential use of this system in modelling climate change scenarios.

\paragraph*{Conclusions:} The project shows that restricted data and
open software can be used in an appropriate way to further the
progress of scientific enquiry.

\vspace{3mm}\hrule
\doublespace

# Background 
## Open software for restricted data

Some datasets such as sensitive personal information about suicide or
climate change scenarios with protected intellectual property need to
be accessed in a restricted way.  In the context of reproducible
research methods, models and assumptions need to be made
transparent and available for scientific debate even though the
datasets may require authorisation to access `r citep(bib[["Peng2011"]])`.

Restrictions around access to data have increased recently in
Australia, especially to the national mortality database after the
discovery of an incident in which Australian population health
researcher Dr Stephen Begg was reported to have 
hacked into information about
deaths in Australia `r citep(bib[["OKeefe2007"]])`.  The subsequent
investigations led to a wide ranging modification to the procedures
for approval and provision of these data that made the access much
more restricted. No new research applications were approved to access these data between 2009 and 2013.

At the same time the reproducibility crisis `r citep(bib[["Peng2011"]])` has emerged,
reducing public confidence in statements by scientists. The true
extent of the problem may turn out to be overstated
`r citep(bib[["Jager2014"]])` however the concern should be addressed
to avoid the problems that a lack of confidence in scientific publications
would entail, especially in respect to evidence-based policy and expenditure of public money. Appropriate access to data and analytic software
addresses this issue.  We investigated available workflow tools for
data management and analysis and implemented a range of these products
on our server.  This server has enhanced our capacity for
experimentation, reviews, revisions and extensions of work in this
field.  We present the results of this project and report that it has
streamlined access to population health and environmental data for
analysis.

## Motivating case study: Drought and suicide

The impact of drought on mental health is plausible, however remains a gap in epidemiological knowledge `r citep(bib[["Stanke2013"]])`.  There is concern too that this health risk will rise under future climate change `r citep(bib[["Berry2008"]])`.   As mentioned the number of studies that have examined the relationship
between suicide and drought is limited to only a handful `r citep(c(bib[["Nicholls2006c"]], bib[["Page2002c"]], bib[["Guiney2012a"]], bib[["Hanigan2012e"]]))`.  The motivating case for this project was to use the historical exposure-response functions to estimate future climate change impacts.

There has been substantial public interest within Australia in recent decades of the putative relationship between drought and rural mental health, including suicide. The topic has frequently been raised by the media, by rural politicians and by mental health support groups `r citep(bib[["AustralianB2006"]])`. There have also been media reports in India indicating substantial concerns about drought and rural suicide in that country `r citep(bib[["SarathiBiswas2012"]])`.

The number of studies that have examined the relationship between suicide and drought is limited. However, many papers explore links between suicide and climate variables other than drought (such as temperature) and there are two major reviews papers available of the literature on climatic influences on suicide `r citep(c(bib[["Dixon2009"]],bib[["Deisenhammer2003b"]]))`. However, very few studies have investigated drought specifically.

There are several mechanisms through which unusually low rainfall, especially if exacerbated by increased soil dryness due to higher temperatures may increase the suicide rate. First, droughts increase the financial stress on farmers and farming communities (even if partially compensated by drought relief welfare payments). Such difficulty may occur in conjunction with other economic stresses, such as rising interest rates, falling commodity prices, or an unfavourable foreign exchange rate.  `r citet(bib[["Vins2015"]])` provide a systematic  literature review of the mental health effects of drought and explore the putative causal mechanisms.

# Methodology

The approach we took to meet the challenge of analysing restricted suicide and climate change scenario data in a safe environment was to
build a new hardware and software stack using open-source software.  We based our
planning on the realisation that there is a growing need of these
technologies in the context of reproducible research.  This
requires that methods, models and assumptions need to be made
transparent and available for scientific debate even though the
datasets may require authorisation to access.  This is true not just
in health data, but also including the context 
of data  with restrictive intellectual property and licence requirements (such as climate change scenario models).

To develop an over-all view of
the issue and analyse the dimensions of the problem we spent the
initial phase of the project conceptualising a rich picture
of the issue, and focused on identifying risks that the project might
face. Several papers that describe similar systems were reviewed `r citep(c(bib[["Evans2012"]], bib[["Fleming2014"]]))` and 
several recommendations from these papers were adopted in our system.

Our design responds directly to the primary threat of unintentional release of
sensitive data so we decided to build a secure server/client environment
for analysts to develop their software in an open way, while ensuring
the safety of the datasets.  Other risks we identified were in
relation to the provision of the server hardware and we were able to
take advantage of the Australian 'Nectar Research Cloud' (\url{http://nectar.org.au}) for virtual machines to build the servers on.

Then we defined the scope and quality of the project outcomes that we
were aiming to deliver.  The fact that restrictions around access to
data have increased recently, coupled with arguments that appropriate
access to analytic software is needed to address the reproducibility
crisis meant that the scope of this project was very broad.  We also
explored the ambitions of our stakeholders to support their publishable
outputs with open software.  Given that examples of un-reproducible work
has spread even to the results published in top journals `r citep(bib[["Peng2015"]])`,
the scope we decided to set for this project was
for very high levels of open-ness for the evidence being presented for
peer-review, along with very high levels of restriction on access to
the data. Luckily however we were able to rule out the need for the
extreme level of restriction such as getting Australian Defence Science and
Technology Organisation (DSTO) accreditation for the security of these
servers against malicious hacking attacks.  Our servers just needed to be tested by the standard 'vulnerability exploits' scanner used by the Australian National University IT Department.

We also looked at the workflow system Kepler to assess it\'s utility
for providing access to the data, but found that there were a lot of
limitations at the time `r citep(bib[["Curcin2008"]])` and decided
that the R environment for statistical computing and graphics would be
the platform we would focus on.

During the next phase of the project we dealt with issues of the costs
associated with developing the software and hosting the hardware at
different locations, as well as the time needed to test and get user
acceptance on the services.

## System design

In this case study we utilise Virtual Machines (VMs) in the cloud. Our
system requires two VMs so that the storage and processing of data can
be compartmentalised, with various benefits.  A high level overview of
the system is shown in Figure \ref{fig:sys}.  Full details including
Linux commands and configuration specifications are available online at
[http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/).

\begin{figure}[!h]
\centering
\includegraphics[width=.85\textwidth]{opensoft.pdf}
\caption{High Level Schematic System Design, colours indicate restrictions (red), open (blue)}
\label{fig:sys}
\end{figure}

## Software selections
We researched a variety of systems and found the following set-up worked best for us.

Linux cluster:

- National Research Cloud [www.nectar.org.au/research-cloud](www.nectar.org.au/research-cloud)
- Centos 6.4 [www.centos.org](www.centos.org)

Geographical Information Systems (GIS) database server:

- PostgreSQL 9.2 [www.postgresql.org](www.postgresql.org)
- PostGIS 2.0 [http://postgis.refractions.net](http://postgis.refractions.net)

Statistical analysis server:

-  R language for statistical computing [www.r-project.org](www.r-project.org)
-  Rstudio server [www.rstudio.com](www.rstudio.com)
-  OpenGeo Suite [http://opengeo.org](http://opengeo.org)

Information management:

- Projects,UsersDB Oracle XE APEX [www.oracle.com](www.oracle.com)
- Data Catalogue [http://assda.anu.edu.au/ddiindex.html](http://assda.anu.edu.au/ddiindex.html)

The client side:

-  Any standard web-browser
-  The Kepler Project [www.kepler-project.org](www.kepler-project.org)
-  pgAdmin [www.pgadmin.org](www.pgadmin.org)
-  Git Version Control and GitHub [www.github.com](www.github.com)
-  Emacs code editor with the starter-kit by Kieran Healy and orgmode [http://kieranhealy.org/resources/emacs-starter-kit/](http://kieranhealy.org/resources/emacs-starter-kit/)

# Case study

## Step 1: Historical exposure-response functions

For this case study we replicated the work we had previously conducted
on our personal desktop computers within the 
university.  A key result is shown in Figure \ref{fig:Figure1.png}.  That
work was published already (Hanigan \emph{et al.} 2012), however
the improved IT infrastructure offered by this project allows the
analysis to be re-run from a secure web-browser interface.  Such
improved access allowed a much broader discussion of the data,
techniques and results because the researcher was able to discuss the
details of the modelling with other scientists at meetings, conferences and
workshops, while actually repeating the computations in real time.
This is a vast improvement over the previous option of leaving the
data analysis on the secure desktop computers at the university, and
merely describing the computations to colleagues at the workshops.

\begin{figure}[!h]
\centering
\includegraphics[width=.45\textwidth]{Figure1.png}
\caption{Drought exposure-response functions Rural Males}
\label{fig:Figure1.png}
\end{figure}

## Step 2: Future drought scenarios and attributable fraction of suicides

Following the methods of `r citet(bib[["Climate2008"]])` we used the climate change scenarios provided for the Garnaut Climate Change Review `r citep(bib[["CSIRO2013"]])` to project estimates of future suicides under various drought conditions.  The statistical method for this calculation is:

$$Y_{ijk}=\sum_{lm}(e^{(\beta_{ijk}} \times { X_{lm})} - 1) \times {BaselineRate_{jkl}} \times { Population_{jklm}}$$

\noindent Where:

$\beta_{ijk}$ = the ExposureVariable coefficient for zone$_i$, age$_j$ and sex$_{k}$

${X_{lm}}$ = Projected Future Exposure Variables

\emph{BaselineRate$_{jkl}$} = \emph{avgDeathsPerTime}/\emph{avgPopPerTime} in age$_j$, sex$_k$ and zone$_l$

\emph{Population$_{jklm}$} = projected populations by age$_j$, sex$_k$, zone$_l$ and time$_m$

```{r, eval = F, echo = F}
dir()
dir("~/data")
indir <- "~/data/GARNAUT mental health archive/1 Data/Rain"
dir(indir, pattern = "jpg")
filelist  <- c("A1BDRY_RainSD07.jpg",             "A1BWET_RainSD07.jpg"            )
for(fi in filelist){
  #file.copy(file.path(indir, fi), fi)
  txt <- sprintf("\\begin{figure}[!h]
\\centering
\\includegraphics[width=\\textwidth]{%s}
\\caption{%s}
\\label{fig:%s}
\\end{figure}
", fi, fi, fi)
  cat(txt)
}
```

## Results

In table \ref{tab:table1} below the two rainfall scenarios used by `r citet(bib[["Berry2008"]])` are used to demonstrate this drought impact assessment report.  For full details of the data from the Garnaut Climate Change Review please refer to the original papers. For the purpose of this discussion only a brief summary is required to appreciate the relevance to this case study. The result shown below merely shows that the estimated impact of climate change can vary a lot given the input datasets and the assumptions and methods applied during analysis.  The codes used to fit models, project scenarios and estimate burden of deaths are all available on the system, and can be assessed by request at the project website: \url{http://swish-climate-impact-assessment.github.com/}

\bigskip

```{r, eval = T, echo = F, results = 'asis'}
library(xtable)
tab <- read.csv(textConnection("Scenario, Deaths per annum, Lower 95th CI, Upper 95th percent CI
Historical (1970-2008), 4.01, 2.14, 6.05
A1FIR1 (Dry), 8.91, 4.56, 14.00
A1FIR2 (Wet), 2.93, 1.5, 4.47
"), strip.white = T)
#tab
names(tab) <- gsub("\\.", " ", names(tab))
print(xtable(tab, caption = 'Estimated number of rural male suicides attributable to drought between 2000 and 2100', label = 'tab:table1'), comment = F, include.rownames = F)

```

\clearpage

Below are pictorial represenations of the climate change scenarios estimated annual total rainfall for each Statistical Division (SD) census region across the country.  SDs are used because this is the geographical unit at which the suicide rates were analysed.

\begin{figure}[!h]
\centering
\includegraphics[width=.5\textwidth]{A1BDRY_RainSD07.jpg}
\caption{A1BDRY RainSD07}
\label{fig:A1BDRYRainSD07}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=.5\textwidth]{A1BWET_RainSD07.jpg}
\caption{A1BWET RainSD07}
\label{fig:A1BWET-RainSD07}
\end{figure}

# Discussion 
## Principal findings

We demonstrate the use of this system by using
the climate change scenarios held on the database to project future
droughts, and use the confidential suicide data that is safely stored
there to estimate baseline risks and future burden of suicide
attributable to the droughts.  Because the server system is easy to
access through a web browser these results are easily reproducible and
alternate scenarios and assumptions can be tested.  Because the webserver is very secure, this work can be done with confidence that the risk of unauthorised data release is minimised.

The results of our project are applicable more generally than just
reproducible research. For instance in relation to issues of
governance and management at the multidisciplinary institutions and
multi-institutional projects it is vitally important that data sharing
is enabled, while some data are still restricted because access requires
approval and authorisation, or because data providers wish to be informed of re-use and identifying the profile of
users of the data.  It is important that data access
can be made restrictive WITHOUT impeding the progress of the local
science agenda (collaborations, workshops, papers etc) and keeping the
relevant data custodian parties informed about what is happening with
the release of their datasets.  The open-ness of the analytical
software also has a positive effect on the value of the data
infrastructure (through education and outreach activities) without
risking any unethical or negligent use of these datasets.

#  Conclusions

In summary we found that:

- Climate data can be attached to suicide and census data in a secure computing environment.
- Future drought associated deaths can be estimated.
- These estimates are very uncertain, contentious and difficult to justify.
- Data management and analysis technology such as that presented is needed to enable rigorous and transparent exploration that challenges or verifies published results.

This system described:

- Enables data analysis in a safe environment.
- Allows comparison of multiple climate scenarios and assumptions.
- Demonstrated with a climate/health impact assessment.
- Addresses the reproducibility crisis.

# References

```{r, echo=FALSE, message=FALSE, eval = T}
write.bibtex(file="references.bib")
```